# Reproducibility

When using the infer package for research, or in other cases when exact reproducibility is a priority, be sure the set the seed for R's random number generator. infer will respect the random seed specified in the `set.seed()` function, returning the same results when `generate()`ing new data given an identical seed. For instance, we can calculate the difference in mean `age` by `college` degree status using the `gss` dataset from 10 versions of the `gss` resampled with permutation using the following code.

```{r, include = FALSE}
library(infer)
```

```{r}
set.seed(1)

gss %>%
  specify(age ~ college) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5, type = "permute") %>%
  calculate("diff in means", order = c("degree", "no degree"))
```

Setting the seed again and rerunning the same code will result in the same answer.

```{r}
# set the seed
set.seed(1)

gss %>%
  specify(age ~ college) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5, type = "permute") %>%
  calculate("diff in means", order = c("degree", "no degree"))
```

Please keep this in mind when writing infer code that utilizes resampling with `generate()`.
