---
title: "Randomization Examples using `nycflights13` `flights` data"
author: "Chester Ismay and Andrew bray"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{flights example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 5) 
```

## Data preparation

```{r message=FALSE, warning=FALSE}
library(nycflights13)
library(dplyr)
library(ggplot2)
library(stringr)
library(infer)
set.seed(2017)
fli_small <- flights %>% 
  sample_n(size = 500) %>% 
  mutate(season = case_when(
    month %in% c(10:12, 1:3) ~ "winter",
    month %in% c(4:9) ~ "summer"
  )) %>% 
  mutate(day_hour = case_when(
    between(hour, 1, 12) ~ "morning",
    between(hour, 13, 24) ~ "not morning"
  )) %>% 
  select(arr_delay, dep_delay, season, 
         day_hour, origin, carrier) %>%
  filter(., complete.cases(.))
```

* Two numeric - `arr_delay`, `dep_delay`
* Two categories 
    - `season` (`"winter"`, `"summer"`), 
    - `day_hour` (`"morning"`, `"not morning"`)
* Three categories - `origin` (`"EWR"`, `"JFK"`, `"LGA"`)
* Sixteen categories - `carrier`

***

## Hypothesis tests

### One numerical variable (mean)

```{r}
x_bar <- fli_small %>%
  summarize(mean(dep_delay)) %>%
  pull()
null <- fli_small %>%
  specify(response = dep_delay) %>%
  hypothesize(null = "point", mu = 10) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean")
ggplot(null, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = x_bar, col = "red")
null %>%
  summarize(mean(stat > x_bar)) * 2
```

### One numerical variable (median): todo


### One categorical (2 level) variable

```{r}
p_hat <- fli_small %>%
  summarize(mean(day_hour == "morning")) %>%
  pull()
null <- fli_small %>%
  specify(response = day_hour, success = "morning") %>%
  hypothesize(null = "point", p = .5) %>% 
  generate(reps = 1000, type = "simulate") %>% 
  calculate(stat = "prop")
ggplot(null, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = p_hat, col = "red")
null %>%
  summarize(mean(stat < p_hat)) * 2
```

### Two categorical (2 level) variables

```{r}
d_hat <- fli_small %>%
  group_by(season) %>%
  summarize(prop = mean(day_hour == "morning")) %>%
  summarize(diff(prop)) %>%
  pull()
null <- fli_small %>%
  specify(day_hour ~ season, success = "morning") %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in props")
ggplot(null, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = d_hat, col = "red")
null %>%
  summarize(mean(stat < d_hat)) * 2
```

### One categorical (>2 level) - GoF

```{r}
null <- fli_small %>%
  specify(response = origin) %>%
  hypothesize(null = "point", p = c("EWR" = .33, "JFK" = .33, "LGA" = .34)) %>% 
  generate(reps = 1000, type = "simulate") %>% 
  calculate(stat = "Chisq")
Chisq_hat <- chisq.test(table(fli_small$origin))$stat
ggplot(null, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = Chisq_hat, col = "red")
null %>%
  summarize(mean(stat > Chisq_hat)) 
```

### Two categorical (>2 level) variables

```{r}
null <- fli_small %>%
  specify(day_hour ~ origin) %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "Chisq")
Chisq_hat <- chisq.test(table(fli_small$day_hour, fli_small$origin))$stat
ggplot(null, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = Chisq_hat, col = "red")
null %>%
  summarize(mean(stat > Chisq_hat)) 
```

One numerical variable, one categorical (2 levels) (diff in means)

```{r}
fli_small %>%
  specify(dep_delay ~ season) %>% # alt: response = age, explanatory = sex
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means") %>% 
  visualize()
```

One numerical variable, one categorical (2 levels) (diff in medians)

```{r}
# prof_small %>%
#   specify(age ~ sex) %>% # alt: response = age, explanatory = sex
#   hypothesize(null = "independence") %>%
#   generate(reps = 1000, type = "permute") %>%
#   calculate(stat = "diff in medians") %>% 
#   visualize()
```

One numerical one categorical (>2 levels) -  ANOVA

```{r}
# prof_small %>%
#   specify(age ~ status) %>% # alt: response = age, explanatory = status
#   hypothesize(null = "independence") %>%
#   generate(reps = 1000, type = "permute") %>%
#   calculate(stat = "F") %>% 
#   visualize()
```

Two numerical vars - SLR 

```{r}
# prof_small %>%
#   specify(age ~ height) %>% # alt: response = age, explanatory = height
#   hypothesize(null = "independence") %>%
#   generate(reps = 1000, type = "permute") %>%
#   calculate(stat = "slope") %>% 
#   visualize()
```

## Confidence intervals

One numerical (one mean)

```{r}
# prof_small %>%
#   specify(response = age) %>%
#   generate(reps = 1000, type = "bootstrap") %>%
#   calculate(stat = "mean") %>% 
#   visualize()
```

One numerical (one median)

```{r}
# prof_small %>%
#   specify(response = age) %>%
#   generate(reps = 1000, type = "bootstrap") %>%
#   calculate(stat = "median") %>% 
#   visualize()
```

One categorical (one proportion)

```{r}
p_hat <- fli_small %>%
  summarize(mean(day_hour == "morning")) %>%
  pull()
boot <- fli_small %>%
  specify(response = day_hour, success = "morning") %>%
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "prop") %>%
  pull()
c(p_hat - 2 * sd(boot), p_hat + 2 * sd(boot))
```

One numerical variable one categorical (2 levels) (diff in means)

```{r}
# prof_small %>%
#   specify(age ~ sex) %>%
#   generate(reps = 1000, type = "bootstrap") %>%
#   calculate(stat = "diff in means") %>% 
#   visualize()
```

Two categorical variables (diff in proportions)

```{r}
d_hat <- fli_small %>%
  group_by(season) %>%
  summarize(prop = mean(day_hour == "morning")) %>%
  summarize(diff(prop)) %>%
  pull()
boot <- fli_small %>%
  specify(day_hour ~ season, success = "morning") %>%
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "diff in props") %>%
  pull()
c(d_hat - 2 * sd(boot), d_hat + 2 * sd(boot))
```

Two numerical vars - SLR

```{r}
# prof_small %>%
#   specify(age ~ height) %>% 
#   generate(reps = 1000, type = "bootstrap") %>%
#   calculate(stat = "slope") %>% 
#   visualize()
```
