---
title: "Randomization Examples using `nycflights13` `flights` data"
author: "Chester Ismay and Andrew Bray"
date: "2018-01-05"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{flights example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

**Note**: The `type` argument in `generate()` is automatically filled based on the entries for `specify()` and
`hypothesize()`. It can be removed throughout the examples that follow. It is left in to reiterate the type of generation process being performed.

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 5) 
```

This vignette is designed to show how to use the {infer} package with {dplyr} syntax. It does not show how to calculate observed statistics or p-values using the {infer} package. To see examples of these, check out the "Computation of observed statistics..." vignette instead.

## Data preparation

```{r message=FALSE, warning=FALSE}
library(nycflights13)
library(dplyr)
library(ggplot2)
library(stringr)
library(infer)
set.seed(2017)
fli_small <- flights %>% 
  na.omit() %>%
  sample_n(size = 500) %>% 
  mutate(season = case_when(
    month %in% c(10:12, 1:3) ~ "winter",
    month %in% c(4:9) ~ "summer"
  )) %>% 
  mutate(day_hour = case_when(
    between(hour, 1, 12) ~ "morning",
    between(hour, 13, 24) ~ "not morning"
  )) %>% 
  select(arr_delay, dep_delay, season, 
         day_hour, origin, carrier)
```

* Two numeric - `arr_delay`, `dep_delay`
* Two categories 
    - `season` (`"winter"`, `"summer"`), 
    - `day_hour` (`"morning"`, `"not morning"`)
* Three categories - `origin` (`"EWR"`, `"JFK"`, `"LGA"`)
* Sixteen categories - `carrier`

***

# Hypothesis tests

### One numerical variable (mean)

```{r}
x_bar <- fli_small %>%
  summarize(mean(dep_delay)) %>%
  pull()
```

```{r}
null_distn <- fli_small %>%
  specify(response = dep_delay) %>%
  hypothesize(null = "point", mu = 10) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean")
ggplot(data = null_distn, mapping = aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = x_bar, color = "red")
null_distn %>%
  summarize(p_value = mean(stat >= x_bar) * 2)
```

### One numerical variable (median)

```{r}
x_tilde <- fli_small %>%
  summarize(median(dep_delay)) %>%
  pull()
null_distn <- fli_small %>%
  specify(response = dep_delay) %>%
  hypothesize(null = "point", med = -1) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "median")
ggplot(null_distn, aes(x = stat)) +
  geom_bar() +
  geom_vline(xintercept = x_tilde, color = "red")
null_distn %>%
  summarize(p_value = mean(stat <= x_tilde) * 2)
```

### One categorical (one proportion)

```{r}
p_hat <- fli_small %>%
  summarize(mean(day_hour == "morning")) %>%
  pull()
null_distn <- fli_small %>%
  specify(response = day_hour, success = "morning") %>%
  hypothesize(null = "point", p = .5) %>%
  generate(reps = 1000, type = "simulate") %>%
  calculate(stat = "prop")
ggplot(null_distn, aes(x = stat)) +
  geom_bar() +
  geom_vline(xintercept = p_hat, color = "red")
null_distn %>%
  summarize(p_value = mean(stat <= p_hat) * 2)
```

Logical variables will be coerced to factors:

```{r}
null_distn <- fli_small %>%
  mutate(day_hour_logical = (day_hour == "morning")) %>%
  specify(response = day_hour_logical, success = "TRUE") %>%
  hypothesize(null = "point", p = .5) %>%
  generate(reps = 1000, type = "simulate") %>%
  calculate(stat = "prop")
```


### Two categorical (2 level) variables

```{r}
d_hat <- fli_small %>%
  group_by(season) %>%
  summarize(prop = mean(day_hour == "morning")) %>%
  summarize(diff(prop)) %>%
  pull()
null_distn <- fli_small %>%
  specify(day_hour ~ season, success = "morning") %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in props", order = c("winter", "summer"))
ggplot(null_distn, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = d_hat, color = "red")
null_distn %>%
  summarize(p_value = mean(stat <= d_hat) * 2) %>%
  pull()
```

### One categorical (>2 level) - GoF

```{r}
Chisq_hat <- fli_small %>%
  specify(response = origin) %>%
  hypothesize(null = "point", 
              p = c("EWR" = .33, "JFK" = .33, "LGA" = .34)) %>% 
  calculate(stat = "Chisq")
null_distn <- fli_small %>%
  specify(response = origin) %>%
  hypothesize(null = "point", 
              p = c("EWR" = .33, "JFK" = .33, "LGA" = .34)) %>% 
  generate(reps = 1000, type = "simulate") %>% 
  calculate(stat = "Chisq")
ggplot(null_distn, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = pull(Chisq_hat), color = "red")
null_distn %>%
  summarize(p_value = mean(stat >= pull(Chisq_hat))) %>%
  pull()
```

### Two categorical (>2 level) variables

```{r}
Chisq_hat <- fli_small %>%
  chisq_stat(formula = day_hour ~ origin)
null_distn <- fli_small %>%
  specify(day_hour ~ origin, success = "morning") %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "Chisq")
ggplot(null_distn, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = Chisq_hat, color = "red")
null_distn %>%
  summarize(p_value = mean(stat >= Chisq_hat)) %>%
  pull()
```

### One numerical variable, one categorical (2 levels) (diff in means)

```{r}
d_hat <- fli_small %>% 
  group_by(season) %>% 
  summarize(mean_stat = mean(dep_delay)) %>% 
  # Since summer - winter
  summarize(-diff(mean_stat)) %>% 
  pull()
null_distn <- fli_small %>%
  specify(dep_delay ~ season) %>% # alt: response = dep_delay, 
  # explanatory = season
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("summer", "winter"))
ggplot(null_distn, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = d_hat, color = "red")
null_distn %>%
  summarize(p_value = mean(stat <= d_hat) * 2) %>%
  pull()
```

### One numerical variable, one categorical (2 levels) (diff in medians)

```{r}
d_hat <- fli_small %>% 
  group_by(season) %>% 
  summarize(median_stat = median(dep_delay)) %>% 
  # Since summer - winter
  summarize(-diff(median_stat)) %>% 
  pull()
null_distn <- fli_small %>%
  specify(dep_delay ~ season) %>% # alt: response = dep_delay, 
  # explanatory = season
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in medians", order = c("summer", "winter"))
ggplot(null_distn, aes(x = stat)) +
  geom_bar() +
  geom_vline(xintercept = d_hat, color = "red")
null_distn %>%
  summarize(p_value = mean(stat >= d_hat) * 2) %>%
  pull()
```

### One numerical, one categorical (>2 levels) -  ANOVA

```{r}
F_hat <- anova(
               aov(formula = arr_delay ~ origin, data = fli_small)
               )$`F value`[1]
null_distn <- fli_small %>%
   specify(arr_delay ~ origin) %>% # alt: response = arr_delay, 
   # explanatory = origin
   hypothesize(null = "independence") %>%
   generate(reps = 1000, type = "permute") %>%
   calculate(stat = "F")
ggplot(null_distn, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = F_hat, color = "red")  
null_distn %>% 
  summarize(p_value = mean(stat >= F_hat)) %>%
  pull()
```

### Two numerical vars - SLR 

```{r}
slope_hat <- lm(arr_delay ~ dep_delay, data = fli_small) %>% 
  broom::tidy() %>% 
  filter(term == "dep_delay") %>% 
  pull(estimate)
null_distn <- fli_small %>%
   specify(arr_delay ~ dep_delay) %>% 
   hypothesize(null = "independence") %>%
   generate(reps = 1000, type = "permute") %>%
   calculate(stat = "slope")
ggplot(null_distn, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = slope_hat, color = "red")  
null_distn %>% 
  summarize(p_value = mean(stat >= slope_hat) * 2) %>%
  pull()
```

## Confidence intervals

### One numerical (one mean)

```{r}
x_bar <- fli_small %>%
   summarize(mean(arr_delay)) %>%
   pull()
boot <- fli_small %>%
   specify(response = arr_delay) %>%
   generate(reps = 1000, type = "bootstrap") %>%
   calculate(stat = "mean") %>%
   pull()
c(lower = x_bar - 2 * sd(boot),
  upper = x_bar + 2 * sd(boot))
```

### One categorical (one proportion)

```{r}
p_hat <- fli_small %>%
 summarize(mean(day_hour == "morning")) %>%
 pull()
boot <- fli_small %>%
 specify(response = day_hour, success = "morning") %>%
 generate(reps = 1000, type = "bootstrap") %>%
 calculate(stat = "prop") %>%
 pull()
c(lower = p_hat - 2 * sd(boot),
 upper = p_hat + 2 * sd(boot))
```

### One numerical variable, one categorical (2 levels) (diff in means)

```{r}
d_hat <- fli_small %>% 
  group_by(season) %>% 
  summarize(mean_stat = mean(arr_delay)) %>% 
  # Since summer - winter
  summarize(-diff(mean_stat)) %>% 
  pull()
boot <- fli_small %>%
   specify(arr_delay ~ season) %>%
   generate(reps = 1000, type = "bootstrap") %>%
   calculate(stat = "diff in means", order = c("summer", "winter")) %>% 
   pull()
c(lower = d_hat - 2 * sd(boot), 
  upper = d_hat + 2 * sd(boot))
```

### Two categorical variables (diff in proportions)

```{r}
d_hat <- fli_small %>%
  group_by(season) %>%
  summarize(prop = mean(day_hour == "morning")) %>%
  # Since summer - winter
  summarize(-diff(prop)) %>%
  pull()
boot <- fli_small %>%
  specify(day_hour ~ season, success = "morning") %>%
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "diff in props", order = c("summer", "winter")) %>%
  pull()
c(lower = d_hat - 2 * sd(boot), 
  upper = d_hat + 2 * sd(boot))
```

### Two numerical vars - SLR

```{r}
slope_hat <- lm(arr_delay ~ dep_delay, data = fli_small) %>% 
  broom::tidy() %>% 
  filter(term == "dep_delay") %>% 
  pull(estimate)
boot <- fli_small %>%
   specify(arr_delay ~ dep_delay) %>% 
   generate(reps = 1000, type = "bootstrap") %>%
   calculate(stat = "slope") %>% 
   pull()
c(lower = slope_hat - 2 * sd(boot), 
  upper = slope_hat + 2 * sd(boot))   
```
