---
title: "infer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{infer}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 3.5) 
options(digits = 4)
```

### Introduction

`infer` implements an expressive grammar to perform statistical inference that coheres with the `tidyverse` design framework. Rather than providing methods for specific statistical tests, this package consolidates the principles that are shared among common hypothesis tests into a set of 4 main verbs (functions,) supplemented with many utilities to visualize and extract value from their outputs.

Regardless of which hypothesis test we're using, we're still asking the same kind of question: is the effect/difference in our observed data real, or due to chance? To answer this question, we start by assuming that the observed data came from some world where "nothing is going on" (i.e. the observed effect was simply due to random chance), and call this assumption our *null hypothesis*. (In reality, we might not believe in the null hypothesis at all---the null hypothesis is in opposition to the *alternate hypothesis*, which supposes that the effect present in the observed data is actually due to the fact that "something is going on.") We then calculate a *test statistic* from our data that describes the observed effect. We can use this test statistic to calculate a *p-value*, giving the probability that our observed data could come about if the null hypothesis was true. If this probability is below some pre-defined *significance level* $\alpha$, then we can reject our null hypothesis.

The workflow of this package is designed around this idea. Starting out with some dataset,

+ `specify()` allows you to specify the variable, or relationship between variables, that you're interested in.
+ `hypothesize()` allows you to declare the null hypothesis.
+ `generate()` allows you to construct a null distribution based on your null hypothesis.
+ `calculate()` allows you to calculate a summary statistic describing the effect observed in the data.

Throughout this vignette, we make use of the *dataset_name* dataset that is included in the package.

### `specify()`: Specifying Response (and Explanatory) Variables

<!-- some background on specify --> 

### `hypothesize()`: Declaring the Null Hypothesis

<!-- some background on hypothesize --> 

### `generate()`: Generating the Null Distribution

<!-- some background on generate --> 

### `calculate()`: Calculating Summary Statistics

<!-- some background on calculate --> 

### Other Utilities in infer

<!-- visualize, get_p_value, get_confidence interval --> 

### Examples

We'll work through some examples using the *dataset_name* dataset. In each, we'll point out the kinds of variable(s) involved, generate a null distribution, calculate the relevant test statistic, visualize where the observed statistic lies in the observed distribution, extract the p-value from the test statistic, and find a confidence interval for the point estimate.

#### Example Template

Point out and describe the relevant variable(s). Mention common terminology that is specific to this use-case.

```{r observed}
# calculate the observed statistic
```

```{r null-dist}
# generate the null distribution
```

```{r test-stat}
# calculate the test statistic
```

```{r visualize}
# visualize where the test statistic lies in the null distribution
```

```{r p-value}
# calculate the p-value from the test statistic
```

```{r confidence-interval}
# find a confidence interval for the point estimate
```




