---
output: github_document
params:
  num_rows: 5000
  rep_times: 5000
---

# Timing analysis

This analysis is for `r params$num_rows` flights with `reps =` ``r params$rep_times``.

```{r}
devtools::session_info()
```


```{r message=FALSE, warning=FALSE}
library(nycflights13)
library(tidyverse)
library(stringr)
library(infer)
set.seed(2017)
fli_small <- flights %>% 
  na.omit() %>% 
  sample_n(size = params$num_rows) %>% 
  mutate(half_year = case_when(
    between(month, 1, 6) ~ "h1",
    between(month, 7, 12) ~ "h2"
  )) %>% 
  mutate(day_hour = case_when(
    between(hour, 1, 12) ~ "morning",
    between(hour, 13, 24) ~ "not morning"
  )) %>% 
  select(arr_delay, dep_delay, half_year, 
         day_hour, origin, carrier)
```

* Two numeric - `arr_delay`, `dep_delay`
* Two categories 
    - `half_year` (`"h1"`, `"h2"`), 
    - `day_hour` (`"morning"`, `"not morning"`)
* Three categories - `origin` (`"EWR"`, `"JFK"`, `"LGA"`)
* Sixteen categories - `carrier`

***

```{r}
library(rbenchmark)
```


One numerical variable (mean)

```{r}
benchmark(
fli_small %>%
  specify(response = arr_delay) %>% # alt: arr_delay ~ NULL (or arr_delay ~ 1)
  hypothesize(null = "point", mu = 50) %>% 
  generate(reps = params$rep_times, type = "bootstrap") %>% 
  calculate(stat = "mean"),
replications = 1, columns = "elapsed"
)
```


One numerical variable (median)

```{r}
benchmark(
fli_small %>%
  specify(response = arr_delay) %>% # alt: arr_delay ~ NULL (or arr_delay ~ 1)
  hypothesize(null = "point", Med = 55) %>%
  generate(reps = params$rep_times, type = "bootstrap") %>%
  calculate(stat = "median"),
replications = 1, columns = "elapsed"
)
```

One categorical (2 level) variable

```{r}
benchmark(
fli_small %>%
  specify(response = half_year) %>% # alt: half_year ~ NULL (or half_year ~ 1)
  hypothesize(null = "point", p = c("h1" = .45)) %>% 
  generate(reps = params$rep_times, type = "simulate") %>% 
  calculate(stat = "prop"),
replications = 1, columns = "elapsed"
)
```

Two categorical (2 level) variables

```{r}
benchmark(
fli_small %>%
  specify(half_year ~ day_hour) %>% # alt: response = half_year, explanatory = vs
  hypothesize(null = "independence") %>%
  generate(reps = params$rep_times, type = "permute") %>%
  calculate(stat = "diff in props"),
replications = 1, columns = "elapsed"
)
```

One categorical (>2 level) - GoF

```{r}
benchmark(
fli_small %>%
  specify(origin ~ NULL) %>% # alt: response = origin
  hypothesize(null = "point", 
              p = c("EWR" = .3, "JFK" = .4, "LGA" = .3)) %>%
  generate(reps = params$rep_times, type = "simulate") %>%
  calculate(stat = "Chisq"),
replications = 1, columns = "elapsed"
)
```

Two categorical (>2 level) variables

```{r}
benchmark(
fli_small %>%
  specify(origin ~ carrier) %>% # alt: response = origin, explanatory = carrier
  hypothesize(null = "independence") %>%
  generate(reps = params$rep_times, type = "permute") %>%
  calculate(stat = "Chisq"),
replications = 1, columns = "elapsed"
)
```

One numerical variable one categorical (2 levels) (diff in means)

```{r}
benchmark(
fli_small %>%
  specify(arr_delay ~ half_year) %>% # alt: response = arr_delay, explanatory = half_year
  hypothesize(null = "independence") %>%
  generate(reps = params$rep_times, type = "permute") %>%
  calculate(stat = "diff in means"),
replications = 1, columns = "elapsed"
)
```

One numerical variable one categorical (2 levels) (diff in medians)

```{r}
benchmark(
fli_small %>%
  specify(arr_delay ~ half_year) %>% # alt: response = arr_delay, explanatory = half_year
  hypothesize(null = "independence") %>%
  generate(reps = params$rep_times, type = "permute") %>%
  calculate(stat = "diff in medians"),
replications = 1, columns = "elapsed"
)
```

One numerical one categorical (>2 levels) -  ANOVA

```{r}
benchmark(
fli_small %>%
  specify(arr_delay ~ origin) %>% # alt: response = arr_delay, explanatory = day_hour
  hypothesize(null = "independence") %>%
  generate(reps = params$rep_times, type = "permute") %>%
  calculate(stat = "F"),
replications = 1, columns = "elapsed"
)
```

Two numerical vars - SLR 

```{r}
benchmark(
fli_small %>%
  specify(arr_delay ~ dep_delay) %>% # alt: response = arr_delay, explanatory = dep_delay
  hypothesize(null = "independence") %>%
  generate(reps = params$rep_times, type = "permute") %>%
  calculate(stat = "slope"),
replications = 1, columns = "elapsed"
)
```

### Confidence intervals

One numerical (one mean)

```{r}
benchmark(
fli_small %>%
  specify(response = arr_delay) %>%
  generate(reps = params$rep_times, type = "bootstrap") %>%
  calculate(stat = "mean"),
replications = 1, columns = "elapsed"
)
```

One numerical (one median)

```{r}
benchmark(
fli_small %>%
  specify(response = arr_delay) %>%
  generate(reps = params$rep_times, type = "bootstrap") %>%
  calculate(stat = "median"),
replications = 1, columns = "elapsed"
)
```

One categorical (one proportion)

```{r}
benchmark(
fli_small %>%
  specify(response = half_year) %>%
  generate(reps = params$rep_times, type = "bootstrap") %>%
  calculate(stat = "prop", success = "h2"),
replications = 1, columns = "elapsed"
)
```

One numerical variable one categorical (2 levels) (diff in means)

```{r}
benchmark(
fli_small %>%
  specify(arr_delay ~ half_year) %>%
  generate(reps = params$rep_times, type = "bootstrap") %>%
  calculate(stat = "diff in means"),
replications = 1, columns = "elapsed"
)
```

Two categorical variables (diff in proportions)

```{r}
benchmark(
fli_small %>%
  specify(half_year ~ day_hour) %>%
  generate(reps = params$rep_times, type = "bootstrap") %>%
  calculate(stat = "diff in props"),
replications = 1, columns = "elapsed"
)
```

Two numerical vars - SLR

```{r}
benchmark(
fli_small %>%
  specify(arr_delay ~ dep_delay) %>% 
  generate(reps = params$rep_times, type = "bootstrap") %>%
  calculate(stat = "slope"),
replications = 1, columns = "elapsed"
)

```
